---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wig

```{r, echo = FALSE, results = "hide", message = FALSE}
# use badges
library(badger)
```

<!-- badges: start -->
![License](https://img.shields.io/badge/license-MIT-blue.svg?style=plastic)
```{r, echo = FALSE, results = "asis"}
cat(
  # badge_doi("10.1016/j.softx.2021.100965", "yellow"),
  badge_lifecycle("stable"),
  # badge_cran_release("rethnicity", "green"),
  # badge_cran_download("rethnicity", "grand-total", "green"),
  # badge_cran_checks("rethnicity"),
  badge_github_actions("fangzhou-xie/rethnicity")
)
```
<!-- badges: end -->

The `wig` package implements the Sinkhorn algorithms for regularized Optimal
Transport problems, Wasserstein Barycenter algorithms for the regularized 
Wasserstein Barycenter problems, Wasserstein Dictionary Learning (WDL) model, 
and **W**asserstein **I**ndex **G**eneration (**WIG**) model 
(see references below).


## Installation

<!-- This package is on [CRAN](https://cran.rstudio.com/), and I recommened  -->
<!-- to use the `pak` to install it: -->

<!-- ``` r -->
<!-- # install pak if not already done so -->
<!-- # install.packages("pak") -->
<!-- pak::pak("wig") -->

<!-- # or you can install it in the classic way -->
<!-- install.packages("wig") -->
<!-- ``` -->

You can install the development version of wig from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("fangzhou-xie/wig")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wig)

## basic example code
```


## Reference

Peyré, G., & Cuturi, M. (2019). Computational Optimal Transport: With Applications to Data Science. Foundations and Trends® in Machine Learning, 11(5–6), 355–607. https://doi.org/10.1561/2200000073

Schmitz, M. A., Heitz, M., Bonneel, N., Ngolè, F., Coeurjolly, D., Cuturi, M., Peyré, G., & Starck, J.-L. (2018). Wasserstein dictionary learning: Optimal transport-based unsupervised nonlinear dictionary learning. SIAM Journal on Imaging Sciences, 11(1), 643–678. https://doi.org/10.1137/17M1140431

Xie, F. (2020). Wasserstein index generation model: Automatic generation of time-series index with application to economic policy uncertainty. Economics Letters, 186, 108874. https://doi.org/10.1016/j.econlet.2019.108874

Xie, F. (2025). Deriving the Gradients of Some Popular Optimal Transport Algorithms (No. arXiv:2504.08722). arXiv. https://doi.org/10.48550/arXiv.2504.08722


<!-- ``` -->
<!-- @article{schmitz2018, -->
<!--   title = {Wasserstein Dictionary Learning: Optimal Transport-Based Unsupervised Nonlinear Dictionary Learning}, -->
<!--   shorttitle = {Wasserstein Dictionary Learning}, -->
<!--   author = {Schmitz, Morgan A. and Heitz, Matthieu and Bonneel, Nicolas and Ngol{\`e}, Fred and Coeurjolly, David and Cuturi, Marco and Peyr{\'e}, Gabriel and Starck, Jean-Luc}, -->
<!--   year = 2018, -->
<!--   month = jan, -->
<!--   journal = {SIAM Journal on Imaging Sciences}, -->
<!--   volume = {11}, -->
<!--   number = {1}, -->
<!--   pages = {643--678}, -->
<!--   issn = {1936-4954}, -->
<!--   doi = {10.1137/17M1140431}, -->
<!--   urldate = {2019-07-15}, -->
<!--   langid = {english} -->
<!-- } -->

<!-- @article{peyre2019, -->
<!--   title = {Computational Optimal Transport: With Applications to Data Science}, -->
<!--   author = {Peyr{\'e}, Gabriel and Cuturi, Marco}, -->
<!--   year = 2019, -->
<!--   month = feb, -->
<!--   journal = {Foundations and Trends\textregistered{} in Machine Learning}, -->
<!--   volume = {11}, -->
<!--   number = {5-6}, -->
<!--   pages = {355--607}, -->
<!--   publisher = {Now Publishers, Inc.}, -->
<!--   issn = {1935-8237, 1935-8245}, -->
<!--   doi = {10.1561/2200000073}, -->
<!--   urldate = {2024-07-29}, -->
<!--   langid = {english}, -->
<!-- } -->

<!-- @article{xie2020, -->
<!--   title = {Wasserstein Index Generation Model: Automatic Generation of Time-Series Index with Application to Economic Policy Uncertainty}, -->
<!--   author = {Xie, Fangzhou}, -->
<!--   year = 2020, -->
<!--   month = jan, -->
<!--   journal = {Economics Letters}, -->
<!--   volume = {186}, -->
<!--   pages = {108874}, -->
<!--   issn = {0165-1765}, -->
<!--   doi = {10.1016/j.econlet.2019.108874}, -->
<!--   urldate = {2019-12-10}, -->
<!-- } -->

<!-- @misc{xie2025, -->
<!--   title = {Deriving the Gradients of Some Popular Optimal Transpor Algorithms}, -->
<!--   author = {Xie, Fangzhou}, -->
<!--   year = 2025, -->
<!--   month = oct, -->
<!--   number = {arXiv:2504.08722}, -->
<!--   eprint = {2504.08722}, -->
<!--   primaryclass = {math}, -->
<!--   publisher = {arXiv}, -->
<!--   doi = {10.48550/arXiv.2504.08722}, -->
<!--   urldate = {2025-11-11}, -->
<!-- } -->
<!-- ``` -->
