---
title: "Sinkhorn Algorithms"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sinkhorn Algorithms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Get Started with Sinkhorn Algorithm

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wig) |> suppressPackageStartupMessages()
```

Suppose we have two probability vectors **a** and **b**, for example,

```{r}
a <- c(.3, .4, .1, .1, .1)
b <- c(.4, .5, .1)
```

and the cost matrix **C** between them

```{r}
C <- rbind(
  c(.1, .2, .3),
  c(.2, .3, .4),
  c(.4, .3, .2),
  c(.3, .2, .1),
  c(.5, .5, .4)
)
reg <- .1
```

and the regularization parameter `reg = .1`.
We can then compute the optimal transport plan by the 
`sinkhorn()` algorithm:

```{r}
sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    verbose = 0
  )
)
```

The optimal coupling matrix (optimal transport plan) is **P**,
and two scaling vectors **f**, **g** with the method being "log".

## Vanilla, Log, and Auto Sinkhorn Algorithms

There are two different methods for the Sinkhorn algorithm implemented
in the `wig` package for the computation of the regularized OT problem:
the vanilla Sinkhorn and the log-stabilized Sinkhorn.

The vanilla algorithm is usually faster, as it is inherently parallel
and the computation can be done purely in 
matrix-vector and matrix-matrix operations.

However, the vanilla algorithm is commonly known for its instability
when the regularization parameter is very small,
and the log-stabilized Sinkhorn will handle any arbitrarily small `reg`.
But the cost of doing so is that the log algorithm is typically slower
than the vanilla one, as it requires column-by-column and row-by-row
soft-minimums.

For example, we can have

```{r}
sol_vanilla <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    method = "vanilla",
    verbose = 0
  )
)
sol_log <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    method = "log",
    verbose = 0
  )
)
```

To speed up the log algorithm, I used multi-threading to speed up the
computation.

```{r}
sol_thread <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    method = "log",
    n_threads = 4,
    verbose = 0
  )
)
```


Moreover, I also provided an "auto" method, trying to provide a heuristics
for determining whether to use "vanilla" or "log" method automatically.
This is also the default algorithm.

Although you mostly likely should avoid doing so,
you can also tweak the `threshold` argument to tune the "auto" method.
The default value is `.1`: smaller means more likely to use "vanilla" method,
bigger means more likely to use "log" method.
If you know which algorithm to use, you should directly set `method`
argument instead of doing this.

```{r}
sol_auto <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    threshold = .1,
    verbose = 0
  )
)
```

See also `vignette("barycenter")`.


## Gradient of Loss w.r.t. **a**

As derived in Xie (2025), I also provide the gradient of the Sinkhorn
algorithm w.r.t. to the source density **a**.
The default for gradient calculation is `FALSE`, 
but you can turn it on by the following:

```{r}
sol_grad <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    with_grad = TRUE,
    verbose = 0
  )
)

sol_grad$grad_a
```

See also `vignette("barycenter")`, `vignette("gradient")`.

## Other Control Parameters

This is the default `sinkhorn_control` argument, 
and we have discussed `reg`, `with_grad`, `threshold` already.

```{r, eval=FALSE}
sinkhorn_control = list(
  reg = .1,
  with_grad = FALSE,
  method = "auto",
  threshold = .1,
  max_iter = 1000L,
  zero_tol = 1e-6,
  verbose = 0L
)
```

The `max_iter` and `zero_tol` arguments are the parameters
to determine maximum number of iterations and the convergence condition
for the Sinkhorn algorithms.

The `verbose` are useful for printing out some useful information
during the computation.
You can set it as an non-negative integer (0 meaning not verbose),
and the algorithm will update info for every `verbose` steps.
For example,

```{r}
sol_verbose <- sinkhorn(
  a,
  b,
  C,
  sinkhorn_control = list(
    reg = reg,
    verbose = 10
  )
)
```

See also `vignette("barycenter")`.


## Reference

Peyré, G., & Cuturi, M. (2019). Computational Optimal Transport: With Applications to Data Science. *Foundations and Trends® in Machine Learning*, 11(5–6), 355–607. https://doi.org/10.1561/2200000073

Xie, F. (2025). Deriving the Gradients of Some Popular Optimal Transport Algorithms (No. arXiv:2504.08722). *arXiv*. https://doi.org/10.48550/arXiv.2504.08722
